---
title: "intestine_enriched_genes"
author: "Rtpw"
date: "3/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages
```{r}
# BiocManager::install("topGO")
```
Load packages
```{r}
library("topGO")
library("tidyverse")
```



# Tissue-specific marker genes analysis

Purpose: evaluate the contamination/enrichment of GFP+ intestine cells by visualizing the log2FoldChange of known tissue-specific genes

```{r}
curated_tissue_genes <- read_csv(file = "../../01_tissue_specific_genes/01_input/Curated_Tissue_Specific_Genes_RTPW.csv") %>% 
  mutate(tissue = fct_relevel(tissue, c("intestine", "germline", "neuron", "muscle", "hypodermis"))) %>%
  mutate(gene_name = fct_rev(fct_reorder(gene_name, as.numeric(tissue))))
res_embryoGFPplus_vs_embryoGFPminus_ashr_df <- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/pairwise_shrunk_DE_results/res_embryoGFPplus_vs_embryoGFPminus_ashr_shrunk.csv")
res_L1GFPplus_vs_L1GFPminus_ashr_df <- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/pairwise_shrunk_DE_results/res_L1GFPplus_vs_L1GFPminus_ashr_shrunk.csv")
res_L3GFPplus_vs_L3GFPminus_ashr_df <- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/pairwise_shrunk_DE_results/res_L3GFPplus_vs_L3GFPminus_ashr_shrunk.csv")
```

```{r}
curated_gene_foldchange <- data.frame(res_embryoGFPplus_vs_embryoGFPminus_ashr_df, stage = "embryo") %>% 
  bind_rows(data.frame(res_L1GFPplus_vs_L1GFPminus_ashr_df, stage = "L1")) %>% 
  bind_rows(data.frame(res_L3GFPplus_vs_L3GFPminus_ashr_df, stage = "L3")) %>% 
  right_join(curated_tissue_genes, by = "WBGeneID") %>%
  mutate(star = case_when(
    padj > 0.01 ~ " ",
    padj < 1*10^-10 ~ "***",
    padj < 1*10^-5 ~ "**",
    padj < 0.01 ~ "*"
                          )) %>%
  ggplot(aes(x = gene_name, y = log2FoldChange, fill = tissue, label = star)) +
  geom_bar(stat = "identity") +
  # geom_hline(yintercept = 1, color = "black", linetype = 2) +
  geom_hline(yintercept = 0, color = "black") +
  facet_wrap(~stage) +
  geom_text() +
  coord_flip()+
  theme_bw()
curated_gene_foldchange
# ggsave(filename = "../03_output/Curated_Gene_Intestine_FoldChange.pdf", plot = curated_gene_foldchange, width = 5, height = 3)
```


# Intestine Enriched Gene Ontology

# Save C. elegans gene ontology table
```{r}
source("../../../David/01_promoters/02_scripts/GOfxns.R")
# paramart <- biomaRt::useMart("parasite_mart", dataset = "wbps_gene", host = "https://parasite.wormbase.org", port = 443)
# WORMGO <- C_elegans_query(paramart)
# 
# saveRDS(WORMGO, file = "../01_input/WORMGO.rds")
# WORMGO<- readRDS(file = "../01_input/WORMGO.rds")
WORMGO_nonObsolete <- read_tsv(file = "../01_input/Celegans_GOterms_NonObsolete_220330.tsv")
WORMGO <- WORMGO_nonObsolete %>% select(wbps_gene_id = "Gene.primaryIdentifier", 
                              external_gene_id = "Gene.symbol",
                              go_accession = "Gene.goAnnotation.ontologyTerm.identifier"
                              )

```

# topGO helper functions

```{r}
mkGOtissue = function(altHyp.df, WORMGO) {
  library(topGO)
  # create a named vector of p-values from DESEQ2 alternative hypothesis method
  allGenes <- altHyp.df %>% drop_na(padj) %>% pull(padj)
  names(allGenes) <- altHyp.df %>% drop_na(padj) %>% pull(WBGeneID)
  # make simple function to determine if a gene has significant p-value or not
  topDiffGenes <- function(geneVec){return(geneVec < 0.01)}
  # assign GO terms to each gene for topGO
  geneID2GO = geneID2GO(WORMGO)
  # set up topGOdata object for each ontology type
  BP.go = new("topGOdata", ontology='BP', 
              allGenes = allGenes,
              geneSel = topDiffGenes,
              nodeSize = 10,
              annot = topGO::annFUN.gene2GO,
              gene2GO = geneID2GO)
  MF.go = new("topGOdata", ontology='MF',
              allGenes = allGenes,
              geneSel = topDiffGenes,
              nodeSize = 10,
              annot = topGO::annFUN.gene2GO,
              gene2GO = geneID2GO)
  CC.go = new("topGOdata", ontology='CC',
              allGenes = allGenes,
              geneSel = topDiffGenes,
              nodeSize = 10,
              annot = topGO::annFUN.gene2GO,
              gene2GO = geneID2GO)
  list(BP=BP.go,CC=CC.go,MF=MF.go)
}

GOSummaryTissue<- function(GOdata, topNodes = 200) {
  library(topGO)
  library(dplyr)
  resultFisher <- topGO::runTest(GOdata, algorithm = "weight01", statistic = "fisher")
  resultKS <- topGO::runTest(GOdata, algorithm = "weight01", statistic = "ks")
  resultFisherParentchild <- topGO::runTest(GOdata, algorithm = "parentchild", statistic = "fisher")
  tab <- topGO::GenTable(
    object=GOdata, 
    ks.pval = resultKS, 
    fisher.pval = resultFisher,
    fisher.PC.pval = resultFisherParentchild,
    orderBy="fisher.pval",
    topNodes = topNodes
  )
  # not sure where the conversion to char is happening. convert back
  # replace ">1e-30" character with number
  tab <- suppressMessages(
    tab %>% mutate(ks.pval = as.numeric(ks.pval), ks.pval = replace_na(ks.pval, 1e-30),
                 fisher.pval = as.numeric(fisher.pval), fisher.pval = replace_na(fisher.pval, 1e-30)
                 )
  )
  return(tab)
}

runGOtissue = function(altHyp.df, WORMGO, topNodes = 200)
{
  go = mkGOtissue(altHyp.df, WORMGO)
  go$BP.result = GOSummaryTissue(go$BP, topNodes)
  go$MF.result = GOSummaryTissue(go$MF, topNodes)
  go$CC.result = GOSummaryTissue(go$CC, topNodes)
  go
}
```

# Embryo intestine GO terms

```{r message=FALSE}
res_embryoGFP_alHyp_greater <- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/res_embryoGFP_alHyp_greater.csv")
embryo_intestine_GO <- runGOtissue(res_embryoGFP_alHyp_greater, WORMGO)
```

# L1 intestine GO terms

```{r message=FALSE}
res_L1GFP_alHyp_greater<- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/res_L1GFP_alHyp_greater.csv")
L1_intestine_GO <- runGOtissue(res_L1GFP_alHyp_greater, WORMGO)
```

# L3 intestine analysis

```{r message=FALSE}
res_L3GFP_alHyp_greater<- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/res_L3GFP_alHyp_greater.csv")
L3_intestine_GO <- runGOtissue(res_L3GFP_alHyp_greater, WORMGO)
```

# topGO plotting function

```{r}
fisherGOplot <- function(in.df){
in.df$BP.result %>% mutate(GO = "BP") %>% bind_rows(in.df$MF.result %>% mutate(GO = "MF")) %>% bind_rows(in.df$CC.result %>% mutate(GO = "CC")) %>%
  filter(Significant > Expected) %>%
  mutate(gene_count = paste0(Significant,"/", round(Expected))) %>%
  slice_min(fisher.pval, n = 10) %>%
  mutate(Term = fct_rev(fct_reorder(Term, fisher.pval))) %>%
ggplot(aes(x = Term, y = -log10(fisher.pval), fill = GO, label = gene_count)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(hjust = -0.05, aes(y = 0))+
  scale_fill_brewer(palette = "Greys", direction = -1) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x=element_text(colour="black"),
        axis.text.y=element_text(colour="black")) +
  ggtitle(paste("data:", deparse(substitute(in.df)), sep = " "))
}

```

```{r fig.width=5, fig.height=3.5}
embryo_intestine_GO_plot <- fisherGOplot(embryo_intestine_GO)
embryo_intestine_GO_plot
# ggsave(plot = embryo_intestine_GO_plot, filename = "../03_output/GO_plots/embryo_intestine_GO_plot.pdf", width = 4.5, height = 5)
```


```{r fig.width=5, fig.height=3.5}
L1_intestine_GO_plot <- fisherGOplot(L1_intestine_GO)
L1_intestine_GO_plot
# ggsave(plot = L1_intestine_GO_plot, filename = "../03_output/GO_plots/L1_intestine_GO_plot.pdf", width = 4.5, height = 5)
```


```{r fig.width=5.25, fig.height=3.5}
L3_intestine_GO_plot <- fisherGOplot(L3_intestine_GO)
L3_intestine_GO_plot
# ggsave(plot = L3_intestine_GO_plot, filename = "../03_output/GO_plots/L3_intestine_GO_plot.pdf", width = 4.5, height = 5)
```
# Intestine enriched per stage upset

```{r fig.width=4, fig.height=3}
embryo_intestine_gene_categories <- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/intestine_gene_categories/embryo_intestine_gene_categories.csv")
L1_intestine_gene_categories <- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/intestine_gene_categories/L1_intestine_gene_categories.csv")
L3_intestine_gene_categories <- read_csv(file = "../../03_emb_L1_L3_intestine_RNAseq/03_output/intestine_gene_categories/L3_intestine_gene_categories.csv")

all_stages_enriched <- embryo_intestine_gene_categories %>% 
  mutate(stage = "embryo") %>% 
  bind_rows(L1_intestine_gene_categories %>% mutate(stage = "L1")) %>% 
  bind_rows(L3_intestine_gene_categories %>% mutate(stage = "L3")) %>% 
  filter(intestine_expression == "enriched") %>%
  group_by(WBGeneID) %>%
  summarise(stages = list(stage))

intestine_upset <- all_stages_enriched %>%
  ggplot(aes(x = stages)) +
  geom_bar(width = 0.5, fill = "black") +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1) +
  scale_x_upset(order_by = "degree") +
  scale_y_continuous(lim = c(0, 1000), name = "# of genes") +
  theme_classic() +
  theme(axis.text.x = element_text(colour = "black"),
        axis.text.y = element_text(colour = "black"))

intestine_upset
# ggsave(intestine_upset, file = "../03_output/Intestine_Enriched_UpSet_Plot.pdf", width = 4, height = 3)
```

```{r}
all_stages_enriched <- all_stages_enriched %>%
  mutate(stage_collapsed = sapply(stages, function(x) paste0(sort(x), collapse = "-")))
```


# Stage specific and stage shared GO terms
```{r}
fisherGOplot <- function(in.df){
in.df$BP.result %>% mutate(GO = "BP") %>% bind_rows(in.df$MF.result %>% mutate(GO = "MF")) %>% bind_rows(in.df$CC.result %>% mutate(GO = "CC")) %>%
  filter(Significant > Expected) %>%
  mutate(gene_count = paste0(Significant,"/", round(Expected))) %>%
  slice_min(fisher, n = 10) %>%
  mutate(Term = fct_rev(fct_reorder(Term, fisher))) %>%
ggplot(aes(x = Term, y = -log10(fisher), fill = GO, label = gene_count)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(hjust = -0.05, aes(y = 0))+
  scale_fill_brewer(palette = "Greys", direction = -1) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x=element_text(colour="black"),
        axis.text.y=element_text(colour="black")) +
  ggtitle(paste("data:", deparse(substitute(in.df)), sep = " "))
}
```


```{r}

embryo_L1_L3_GO <- runGO(foreground_genes = all_stages_enriched %>% filter(stage_collapsed == "embryo-L1-L3") %>% pull(WBGeneID),
     background_genes = res_embryoGFP_alHyp_greater %>% pull(WBGeneID),
    WORMGO = WORMGO,
    topNodes = 20)
fisherGOplot(embryo_L1_L3_GO)
```

```{r}
embryo_only_GO <- runGO(foreground_genes = all_stages_enriched %>% filter(stage_collapsed == "embryo") %>% pull(WBGeneID),
     background_genes = res_embryoGFP_alHyp_greater %>% pull(WBGeneID),
    WORMGO = WORMGO,
    topNodes = 20)
fisherGOplot(embryo_only_GO)
```
```{r}
L1_only_GO <- runGO(foreground_genes = all_stages_enriched %>% filter(stage_collapsed == "L1") %>% pull(WBGeneID),
     background_genes = res_embryoGFP_alHyp_greater %>% pull(WBGeneID),
    WORMGO = WORMGO,
    topNodes = 20)
fisherGOplot(L1_only_GO)
```

```{r}
L3_only_GO <- runGO(foreground_genes = all_stages_enriched %>% filter(stage_collapsed == "L3") %>% pull(WBGeneID),
     background_genes = res_embryoGFP_alHyp_greater %>% pull(WBGeneID),
    WORMGO = WORMGO,
    topNodes = 20)
fisherGOplot(L3_only_GO)
```
```{r}
L1_L3_shared_GO <- runGO(foreground_genes = all_stages_enriched %>% filter(stage_collapsed == "L1-L3") %>% pull(WBGeneID),
     background_genes = res_embryoGFP_alHyp_greater %>% pull(WBGeneID),
    WORMGO = WORMGO,
    topNodes = 20)
fisherGOplot(L1_L3_shared_GO)
```
```{r}
embryo_L1_shared_GO <- runGO(foreground_genes = all_stages_enriched %>% filter(stage_collapsed == "embryo-L1") %>% pull(WBGeneID),
     background_genes = res_embryoGFP_alHyp_greater %>% pull(WBGeneID),
    WORMGO = WORMGO,
    topNodes = 20)
fisherGOplot(embryo_L1_shared_GO)
```
```{r}
embryo_L3_shared_GO <- runGO(foreground_genes = all_stages_enriched %>% filter(stage_collapsed == "embryo-L3") %>% pull(WBGeneID),
     background_genes = res_embryoGFP_alHyp_greater %>% pull(WBGeneID),
    WORMGO = WORMGO,
    topNodes = 20)
fisherGOplot(embryo_L3_shared_GO)

embryo_L3_shared_GO$BP.result %>% mutate(ES = Significant/Expected) %>% arrange(desc(ES))

embryo_L3_shared_GO$MF.result
genesInTerm(embryo_L3_shared_GO$MF)[["GO:0016491"]][genesInTerm(embryo_L3_shared_GO$MF)[["GO:0016491"]] %in% (all_stages_enriched %>% filter(stage_collapsed == "embryo-L3") %>% pull(WBGeneID))]
```
<!-- # GO graph structure -->

<!-- ```{r} -->
<!-- # BiocManager::install("Rgraphviz") -->
<!-- library("Rgraphviz") -->

<!-- emb_go_BP_res_weight_fisher <- runTest(embryo_intestine_GO$BP, algorithm = "weight01", statistic = "fisher") -->
<!-- GenTable(embryo_intestine_GO$BP, pval = emb_go_BP_res_weight_fisher) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- par(cex = 0.25) -->
<!-- showSigOfNodes(embryo_intestine_GO$BP, score(emb_go_BP_res_weight_fisher), firstSigNodes = 5, useInfo = "all") -->

<!-- ``` -->
<!-- Parentchild agolrithm returns terms that are too general -->

<!-- ```{r} -->
<!-- emb_go_BP_res_parentchild_fisher <- runTest(embryo_intestine_GO$BP, algorithm = "parentchild", statistic = "fisher") -->
<!-- GenTable(embryo_intestine_GO$BP, pval = emb_go_BP_res_parentchild_fisher) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- par(cex = 0.25) -->
<!-- showSigOfNodes(embryo_intestine_GO$BP, score(emb_go_BP_res_parentchild_fisher), firstSigNodes = 5, useInfo = "all") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- L1_go_BP_res_weight_fisher <- runTest(L1_intestine_GO$BP, algorithm = "weight01", statistic = "fisher") -->
<!-- GenTable(L1_intestine_GO$BP, pval = L1_go_BP_res_weight_fisher) -->
<!-- par(cex = 0.25) -->
<!-- showSigOfNodes(L1_intestine_GO$BP, score(L1_go_BP_res_weight_fisher), firstSigNodes = 5, useInfo = "all", ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- L3_go_BP_res_weight_fisher <- runTest(L3_intestine_GO$BP, algorithm = "weight01", statistic = "fisher") -->
<!-- GenTable(L3_intestine_GO$BP, pval = L3_go_BP_res_weight_fisher) -->
<!-- par(cex = 0.25) -->
<!-- showSigOfNodes(L3_intestine_GO$BP, score(L3_go_BP_res_weight_fisher), firstSigNodes = 5, useInfo = "all", ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- emb_go_MF_res_weight_fisher <- runTest(embryo_intestine_GO$MF, algorithm = "weight01", statistic = "fisher") -->
<!-- GenTable(embryo_intestine_GO$MF, pval = emb_go_MF_res_weight_fisher) -->
<!-- par(cex = 0.25) -->
<!-- showSigOfNodes(embryo_intestine_GO$MF, score(emb_go_MF_res_weight_fisher), firstSigNodes = 5, useInfo = "all") -->
<!-- ``` -->

