---
title: "elt2_target_gene_ontology"
author: "Robert Williams"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library("topGO")
library(tidyverse)
```
# Gene ontology analysis

```{r}
source("../../../David/01_promoters/02_scripts/GOfxns.R")

WORMGO_nonObsolete <- read_tsv(file = "../../07_intestine_enriched_genes/01_input/Celegans_GOterms_NonObsolete_220330.tsv")
WORMGO <- WORMGO_nonObsolete %>% dplyr::select(wbps_gene_id = "Gene.primaryIdentifier", 
                              external_gene_id = "Gene.symbol",
                              go_accession = "Gene.goAnnotation.ontologyTerm.identifier"
                              )


fisherGOplot <- function(in.df){
in.df$BP.result %>% mutate(GO = "BP") %>% bind_rows(in.df$MF.result %>% mutate(GO = "MF")) %>% bind_rows(in.df$CC.result %>% mutate(GO = "CC")) %>%
  filter(Significant > Expected) %>%
  mutate(gene_count = paste0(Significant,"/", round(Expected))) %>%
  slice_min(fisher, n = 10) %>%
  mutate(Term = fct_rev(fct_reorder(Term, fisher))) %>%
ggplot(aes(x = Term, y = -log10(fisher), fill = GO, label = gene_count)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(hjust = -0.05, aes(y = 0))+
  scale_fill_brewer(palette = "Greys", direction = -1) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x=element_text(colour="black"),
        axis.text.y=element_text(colour="black")) +
  ggtitle(paste("data:", deparse(substitute(in.df)), sep = " "))
}
```

```{r}
embryo_activated_GO <- runGO(foreground_genes = all_stages_chip %>% filter(stage == "embryo", elt2_ko == "activated", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
fisherGOplot(embryo_activated_GO)
```
```{r}
embryo_repressed_GO <- runGO(foreground_genes = all_stages_chip %>% filter(stage == "embryo", elt2_ko == "repressed", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
fisherGOplot(embryo_repressed_GO)
```
```{r}
L1_activated_GO <- runGO(foreground_genes = all_stages_chip %>% filter(stage == "L1", elt2_ko == "activated", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
fisherGOplot(L1_activated_GO)
```

```{r}
L1_repressed_GO <- runGO(foreground_genes = all_stages_chip %>% filter(stage == "L1", elt2_ko == "repressed", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
fisherGOplot(L1_repressed_GO)
```

```{r}
L3_activated_GO <- runGO(foreground_genes = all_stages_chip %>% filter(stage == "L3", elt2_ko == "activated", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
fisherGOplot(L3_activated_GO)
```

```{r}
L3_repressed_GO <- runGO(foreground_genes = all_stages_chip %>% filter(stage == "L3", elt2_ko == "repressed", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
fisherGOplot(L3_repressed_GO)
```
# What terms are ELT-2 activated genes enriched for regardless of developmental stage?

```{r}
all_enriched_activated_targets_GO <- runGO(foreground_genes = all_stages_chip %>% filter(intestine_expression == "enriched",elt2_ko == "activated", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
all_enriched_activated_targets_GO_plot <- fisherGOplot(all_enriched_activated_targets_GO)
all_enriched_activated_targets_GO_plot
ggsave(filename = "../03_output/GO_plots/all_enriched_activated_targets_GO_plot.pdf", plot = all_enriched_activated_targets_GO_plot, width = 4, height = 5)
```

```{r}
all_not_enriched_activated_targets_GO <- runGO(foreground_genes = all_stages_chip %>% filter(intestine_expression == "not_enriched",elt2_ko == "activated", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
all_not_enriched_activated_targets_GO_plot <- fisherGOplot(all_not_enriched_activated_targets_GO)
all_not_enriched_activated_targets_GO_plot
ggsave(filename = "../03_output/GO_plots/all_not_enriched_activated_targets_GO_plot.pdf", plot = all_not_enriched_activated_targets_GO_plot, width = 4, height = 5)
```

# What terms are ELT-2 repressed genes enriched for regardless of developmental stage?

```{r}
all_enriched_repressed_targets_GO <- runGO(foreground_genes = all_stages_chip %>% filter(intestine_expression == "enriched",elt2_ko == "repressed", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
all_enriched_repressed_targets_GO_plot <- fisherGOplot(all_enriched_repressed_targets_GO)
all_enriched_repressed_targets_GO_plot
ggsave(filename = "../03_output/GO_plots/all_enriched_repressed_targets_GO_plot.pdf", plot = all_enriched_repressed_targets_GO_plot, width = 4, height = 5)
```

```{r}
all_not_enriched_repressed_targets_GO <- runGO(foreground_genes = all_stages_chip %>% filter(intestine_expression == "not_enriched",elt2_ko == "repressed", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
all_not_enriched_repressed_targets_GO_plot <- fisherGOplot(all_not_enriched_repressed_targets_GO)
all_not_enriched_repressed_targets_GO_plot
ggsave(filename = "../03_output/GO_plots/all_not_enriched_repressed_targets_GO_plot.pdf", plot = all_not_enriched_repressed_targets_GO_plot, width = 4, height = 5)
```

# What terms are ELT-2 activated target genes enriched for regardless of developmental stage or intestine expression?

```{r}
all_activated_targets_GO <- runGO(foreground_genes = all_stages_chip %>% filter(elt2_ko == "activated", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
all_activated_targets_GO_plot <- fisherGOplot(all_activated_targets_GO)
all_activated_targets_GO_plot
ggsave(filename = "../03_output/GO_plots/all_activated_targets_GO_plot.pdf", plot = all_activated_targets_GO_plot, width = 4, height = 5)
```

# What terms are ELT-2 target repressed genes enriched for regardless of developmental stage or intestine expression?

```{r}
all_repressed_targets_GO <- runGO(foreground_genes = all_stages_chip %>% filter(elt2_ko == "repressed", promoter_status == "bound") %>% pull(WBGeneID),
                             background_genes = all_stages_chip %>% pull(WBGeneID) %>% unique(),
                            WORMGO = WORMGO,
                            topNodes = 20)
all_repressed_targets_GO_plot <- fisherGOplot(all_repressed_targets_GO)
all_repressed_targets_GO_plot
ggsave(filename = "../03_output/GO_plots/all_repressed_targets_GO_plot.pdf", plot = all_repressed_targets_GO_plot, width = 4, height = 5)
```