---
title: "embryo_L1_L3_intestine_RNAseq"
author: "Rtpw"
date: "2/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(DESeq2)
library(corrplot)
library(RColorBrewer)
library(pheatmap)
library(tidyverse)
library(apeglm)
library(ashr)
library(InterMineR)
library(ComplexHeatmap)
```

```{r source-functions}
source(file = "../../elt2_paper_functions.R")
```


# Analysis outline

- Import counts and format for DESeq2 analysis
- Filter low abundance counts based on counts per million (CPM)
- Perform differential expression analysis
- Visualize pairwise sample correlation
- Categorize gene expression with alternative hypothesis
- Visualize unshrunken pairwise differential expression
- Visualize shrunken pairwise differential expression
- Measure tissue expressed genes
- Visualize developmentally dynamic intestine genes
- Identify novel intestine genes in each stage for smFISH experiments

# Import and format count data

```{r}
# import counts
countsData <- read.delim(file = "../01_input/all.counts", sep = " ")
# preview counts
head(countsData)
# print samples
colnames(countsData[6:ncol(countsData)])
# import metadata and process file
metadata1 <- read.table(file = "../01_input/RWP27_metadata.tsv", header = FALSE, stringsAsFactors = FALSE) %>% bind_rows(read.table(file = "../01_input/RWP26_metadata.tsv", header = FALSE, stringsAsFactors = FALSE)) %>%
  bind_rows(read.table(file = "../01_input/RWP30_metadata.tsv", header = FALSE, stringsAsFactors = FALSE))

colnames(metadata1) <- c("Filename.Fwd", "Filename.Rev", "names")
head(metadata1)

# separate and process sample info
metadata1 <- metadata1 %>% separate(names, sep = "_", into = c("stage", "sample", "rep"), remove = FALSE)
metadata1 <- metadata1 %>% mutate(stage = fct_relevel(stage, c("embryo", "L1", "L3")), 
                     sample = fct_relevel(sample, c("whole", "cells", "GFPplus", "GFPminus")),
                     rep = fct_relevel(rep, c("rep1", "rep2", "rep3")),
                     names = fct_relevel(names, metadata1$names)
                     )

# Order columns according to metadata1 order
countsData <- countsData  %>% select(chr:length, sort(metadata1$names))
head(countsData)

# Generate a table called "cts" out of the countsData table. Subset the countsData.
cts <- as.matrix(countsData %>% select(metadata1$names))
head(cts)

# Reorganize the metadata table so the names2 column are now headers
rownames(metadata1)<- metadata1$names
coldata <- metadata1[,c("names", "stage", "sample", "rep")]
rownames(coldata) <- as.vector(metadata1$names)
# make grouping variable
coldata$group <- factor(paste0(coldata$stage, coldata$sample))
coldata

# Check that the names match  --> Should be TRUE
all(rownames(coldata) == colnames(cts))

```


# Make DESeqDataSet

Generate the DESeqDataSet. The variables in this design formula will be the type of sample, and the preparation date. This should reduce the variability between the samples based on when they were made.

From the vignette: "In order to benefit from the default settings of the package, you should put the variable of interest at the end of the formula and make sure the control level is the first level."

The variable of interest is the sample `type`.

Using `DESeqDataSetFromMatrix` since I used the program `featureCounts`.

```{r}
dds <- DESeqDataSetFromMatrix(countData = cts,
                              colData = coldata,
                              design = ~ group)
```

Visualize read count distribution
```{r}
raw_count_threshold <- 10
hist(log10(rowSums(counts(dds))), breaks = 50)
abline(v = log10(raw_count_threshold), col = "red", lty = 2)
```
# Filter genes with sum counts per million >= 10 across all samples

```{r}
cpm <- apply(counts(dds),2, function(x) (x/sum(x))*1000000)
hist(log10(rowSums(cpm)))
abline(v = log10(raw_count_threshold), col = "red", lty = 2)
```

Filter genes with low read counts
```{r}
keep <- rowSums(cpm) >= raw_count_threshold
dds <- dds[keep,]
dds
```

Perform Differential Expression
```{r}
dds <- DESeq(dds)
resultsNames(dds)
```

# Sample-to-sample distance matrix

```{r}
vsd <- vst(dds, blind = FALSE)
sampleDists <- dist(t(assay(vsd)))
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$names
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
mega_cor_plot <- pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors)
mega_cor_plot
```
```{r}
myPDFplot(plot = mega_cor_plot, name = "FACS_Correlation_All_Samples", height = 4.5, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")

```

```{r}
all_embryo_correlation <- vsd.corr.per.stage("embryo", "Embryo Stage Intestine FACS RNA-seq Correlation")
all_embryo_correlation
myPDFplot(plot = all_embryo_correlation, name = "Embryo_FACS_Correlation_All_Samples", height = 4, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")
```


```{r}
all_L1_correlation <- vsd.corr.per.stage("L1", "L1 Stage Intestine FACS RNA-seq Correlation")
all_L1_correlation
myPDFplot(plot = all_L1_correlation, name = "L1_FACS_Correlation_All_Samples", height = 4, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")
```


```{r}
all_L3_correlation <- vsd.corr.per.stage("L3", "L3 Stage Intestine FACS RNA-seq Correlation")
myPDFplot(plot = all_L3_correlation, name = "L3_FACS_Correlation_All_Samples", height = 4, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")
```
```{r}
all_sorted_samples_cor <- vsd.corr.per.stage("GFPplus|GFPminus", "Correlation of FACS isolated GFP+ and GFP- samples")
myPDFplot(plot = all_sorted_samples_cor, name = "All_Stage_FACS_Correlation_Sorted_Samples", height = 4, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")
```
Per-stage GFP+ and GFP- correlation
```{r}
embryo_sortedOnly_correlation <- vsd.corr.per.stage("embryo_GFPplus|embryo_GFPminus", "Correlation of Embryo FACS isolated GFP+ and GFP- samples")
myPDFplot(plot = embryo_sortedOnly_correlation, name = "Embryo_FACS_Correlation_Sorted_Only", height = 4, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")
```
```{r}
L1_sortedOnly_correlation <- vsd.corr.per.stage("L1_GFPplus|L1_GFPminus", "Correlation of L1 FACS isolated GFP+ and GFP- samples")
myPDFplot(plot = L1_sortedOnly_correlation, name = "L1_FACS_Correlation_Sorted_Only", height = 4, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")
```
```{r}
L3_sortedOnly_correlation <- vsd.corr.per.stage("L3_GFPplus|L3_GFPminus", "Correlation of L3 FACS isolated GFP+ and GFP- samples")
myPDFplot(plot = L3_sortedOnly_correlation, name = "L3_FACS_Correlation_Sorted_Only", height = 4, width = 6, plotdir = "../03_output/plots/Correlation_Matrix/")
```
# Remove L1 rep 2

```{r}
remove_samples <- c("L1_whole_rep2", "L1_cells_rep2", "L1_GFPplus_rep2", "L1_GFPminus_rep2")
coldata <- coldata %>% filter(!names %in% remove_samples)
dds <- dds[,!colnames(dds)%in% remove_samples]
```

# Perform Differential Expression
```{r}
dds <- DESeq(dds)
resultsNames(dds)
```

# Within-stage pairwise comparisons

<!-- ## Code testing, preliminary visualization -->

<!-- ```{r} -->
<!-- #for loop attempt one -->
<!-- combinations<- data.frame(combos = 1:4, top = c("embryowhole", "embryoGFPplus", "embryoGFPminus", "embryoGFPplus"), bottom = c("embryocells", "embryocells", "embryocells", "embryoGFPminus")) -->

<!-- res_embryowhole_vs_embryocells <- results(dds, contrast = c("group", combinations$top[1], combinations$bottom[1])) -->
<!-- res_embryowhole_vs_embryocells_shrunk <- lfcShrink(dds, contrast = c("group", combinations$top[1], combinations$bottom[1]), type = "ashr", quiet = TRUE) -->

<!-- res_embryowhole_vs_embryocells <- results(dds, contrast = c("group", combinations$top[1], combinations$bottom[1])) -->
<!-- res_embryowhole_vs_embryocells_shrunk <- lfcShrink(dds, contrast = c("group", combinations$top[1], combinations$bottom[1]), type = "ashr", quiet = TRUE) -->

<!-- for(i in 1:nrow(combinations)){ -->
<!--   res.var.name<- paste("res",combinations$top[i], "vs", combinations$bottom[i], sep = "_") -->
<!--   res.shrunk.var.name <- paste("res",combinations$top[i], "vs", combinations$bottom[i], "shrunk", sep = "_") -->
<!--   print(res.var.name) -->
<!--   print(res.shrunk.var.name) -->
<!--   assign(res.var.name, results(dds, contrast = c("group", combinations$top[i], combinations$bottom[i]))) -->
<!-- } -->

<!-- # res_embryoGFPplus_vs_embryoGFPminus -->
<!-- # plotMA(res_embryowhole_vs_embryocells_shrunk, ylim = ylim, main = "whole vs cells") -->
<!-- # combinations$top[1] -->
<!-- # class(res_embryowhole_vs_embryocells) -->

<!-- # for loop attempt two -->
<!-- samples <- paste("embryo", c("whole", "GFPplus", "cells", "GFPminus"), sep = "") -->
<!-- combos <- combn(samples, 2, simplify = FALSE) -->

<!-- for(i in 1:length(combos)){ -->
<!--   print(c(combos[[i]][1], combos[[i]][2])) -->
<!--   res.var.name<- paste("res",combos[[i]][1], "vs", combos[[i]][2], sep = "_") -->
<!--   res.shrunk.var.name <- paste("res",combos[[i]][1], "vs", combos[[i]][2], "shrunk", sep = "_") -->
<!--   print(res.var.name) -->
<!--   print(res.shrunk.var.name) -->
<!--   assign(res.var.name, results(dds, contrast = c("group", combos[[i]][1], combos[[i]][2]))) -->
<!--   assign(res.shrunk.var.name, lfcShrink(dds, contrast = c("group", combos[[i]][1], combos[[i]][2]), type = "ashr", quiet = TRUE)) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- combos -->
<!-- ``` -->


<!-- ```{r} -->
<!-- par(mfrow = c(3, 3), mar=c(2,2,1,1)) -->
<!-- ylim = c(-10,10) -->
<!-- plotMA(res_embryowhole_vs_embryocells_shrunk, ylim = ylim, main = "whole vs cells") -->
<!-- plotMA(res_embryowhole_vs_embryoGFPplus_shrunk, ylim = ylim, main = "whole vs GFP+") -->
<!-- plotMA(res_embryowhole_vs_embryoGFPminus_shrunk, ylim = ylim, main = "whole vs GFP-") -->
<!-- plotMA(res_embryoGFPplus_vs_embryocells_shrunk, ylim = ylim, main = "GFP+ vs cells") -->
<!-- plotMA(res_embryoGFPplus_vs_embryoGFPminus_shrunk, ylim = ylim, main = "GFP+ vs GFP-") -->
<!-- plotMA(res_embryocells_vs_embryoGFPminus_shrunk, ylim = ylim, main = "cells vs GFP-") -->
<!-- ``` -->
<!-- ## ggplot version of plotting -->

<!-- ```{r} -->
<!-- all_embryo_comparisons <- data.frame() -->
<!-- for(i in 1:length(combos)){ -->
<!-- tobind <- as.data.frame(lfcShrink(dds, contrast = c("group", combos[[i]][1], combos[[i]][2]), type = "ashr", quiet = TRUE)) %>%  -->
<!--   rownames_to_column(var = "WBGeneID") %>%  -->
<!--   mutate(comparison = paste(combos[[i]][1], combos[[i]][2], sep = "_vs_"))  -->
<!-- all_embryo_comparisons <- bind_rows(all_embryo_comparisons, tobind) -->
<!-- } -->


<!-- ggplot(all_embryo_comparisons %>% replace_na(list(padj = 1)), aes(x = log10(baseMean), y = log2FoldChange, color = padj < 0.1)) + -->
<!--   geom_point(shape = 16, alpha = 0.1, stroke = 0, size = 1) + -->
<!--   ylim(c(-10,10))+ -->
<!--   facet_wrap(~comparison) + -->
<!--   scale_color_manual(values = c("black", "red"), name = "q.value < 0.1") + -->
<!--   theme_classic() -->


<!-- # write_csv(all_embryo_comparisons, file = "../03_output/embryo_all_pairwise_res.csv") -->
<!-- ``` -->

## Set cutoff values
```{r}
thresh = 1
sig = 0.01
```


```{r}
embryo_alt_hyp_res_df <- alt_hyp_res_df("embryo", thresh = thresh, sig = sig)

# embryo_alt_hyp_res_df %>% replace_na(list(padj = 1)) %>% pull(baseMean) %>% range
```


```{r}
de_category_MA_plot(embryo_alt_hyp_res_df, paste("Embryo differentially expressed genes\nlfc = ",thresh," & padj < ",sig, sep = ""))
```
```{r}
de_category_bar_plot(embryo_alt_hyp_res_df, paste("Embryo differentially expressed genes\nlfc = ",thresh," & padj < ",sig, sep = ""))
```

```{r}
L1_alt_hyp_res_df<- alt_hyp_res_df("L1", thresh = thresh, sig = sig)
```

```{r}
de_category_MA_plot(L1_alt_hyp_res_df, paste("L1 differentially expressed genes\nlfc = ",thresh," & padj < ",sig, sep = ""))
```
```{r}
de_category_bar_plot(L1_alt_hyp_res_df, paste("L1 differentially expressed genes\nlfc = ",thresh," & padj < ",sig, sep = ""))
```


```{r}
L3_alt_hyp_res_df<- alt_hyp_res_df("L3", thresh = thresh, sig = sig)
```

```{r}
de_category_MA_plot(L3_alt_hyp_res_df, paste("L3 differentially expressed genes\nlfc = ",thresh," & padj < ",sig, sep = ""))
```
```{r}
de_category_bar_plot(L3_alt_hyp_res_df, paste("L3 differentially expressed genes\nlfc = ",thresh," & padj < ",sig, sep = ""))
```



# Embryo Intestine FACS MA plots
```{r}
embryo_pairwise_res_shrunk <- shrunk_pairwise_array_df(stage = "embryo")

MA_plot_array(embryo_pairwise_res_shrunk, "embryo FACS ashr shrunk", sig = 0.01)
```

# Label differentially expressed genes on the lfcShrunk plots
Label the shrunken plots with expression status.
doesn't work, currently giving error:
`Error: vector memory exhausted (limit reached?)`

```{r}
embryo_pairwise_res_shrunk_DEtype <- left_join(embryo_pairwise_res_shrunk, embryo_alt_hyp_res_df %>% select(WBGeneID, type, label, isDE), by = c("WBGeneID","label"))
MA_plot_array
ggplot(embryo_pairwise_res_shrunk_DEtype %>% filter(isDE == TRUE), aes(x = log10(baseMean), y = log2FoldChange, color = type)) +
  geom_point(data = embryo_pairwise_res_shrunk_DEtype %>% mutate(padj = replace_na(padj, 1)), shape = 16, alpha = 0.01, stroke = 0, size = 1, color = "grey") +
  geom_point(shape = 16, alpha = 0.1, stroke = 0, size = 1) +
  ylim(c(-10,10))+
  facet_wrap(~label) +
  # scale_color_manual(values = c("black", "red"), name = "q.value < 0.1") +
  theme_classic()

table(embryo_pairwise_res_shrunk_DEtype$type, embryo_pairwise_res_shrunk_DEtype$isDE)
```

# L1 Intestine FACS MA plots
```{r}
L1_pairwise_res_shrunk <- shrunk_pairwise_array_df(stage = "L1")

MA_plot_array(L1_pairwise_res_shrunk, "L1 FACS ashr shrunk", sig = 0.01)
```

# L3 Intestine FACS MA plots
```{r}
L3_pairwise_res_shrunk <- shrunk_pairwise_array_df(stage = "L3")
MA_plot_array(L3_pairwise_res_shrunk, "L3 FACS ashr shrunk", sig = 0.01)
```

# DESeq2 builtin plotting function
```{r}
res_embryoGFPplus_vs_embryoGFPminus <- results(dds, contrast = c("group", "embryoGFPplus", "embryoGFPminus"))
res_L1GFPplus_vs_L1_GFPminus <- results(dds, contrast = c("group", "L1GFPplus", "L1GFPminus"))
res_L3GFPplus_vs_L3_GFPminus <- results(dds, contrast = c("group", "L3GFPplus", "L3GFPminus"))
```

```{r fig.width=6, fig.height=2}
par(mfrow=c(1,3),mar=c(2,2,1,1))
ylim <- c(-15,15)
# drawLines <- function() abline(h=c(-2,2),col="dodgerblue",lwd=2)
sig = 0.01
plotMA(res_embryoGFPplus_vs_embryoGFPminus, ylim=ylim, main = "Embryo GFP+ vs GFP-", alpha = sig)
plotMA(res_L1GFPplus_vs_L1_GFPminus, ylim=ylim, main = "L1 GFP+ vs GFP-", alpha = sig)
plotMA(res_L3GFPplus_vs_L3_GFPminus, ylim=ylim, main = "L3 GFP+ vs GFP-", alpha = sig)
```

```{r}
res_embryoGFPplus_vs_embryoGFPminus_ashr <- lfcShrink(dds, contrast = c("group", "embryoGFPplus", "embryoGFPminus"), type = "ashr")
res_L1GFPplus_vs_L1GFPminus_ashr <- lfcShrink(dds, contrast = c("group", "L1GFPplus", "L1GFPminus"), type = "ashr")
res_L3GFPplus_vs_L3GFPminus_ashr <- lfcShrink(dds, contrast = c("group", "L3GFPplus", "L3GFPminus"), type = "ashr")


# write_csv(as.data.frame(res_embryoGFPplus_vs_embryoGFPminus_apeglm) %>% rownames_to_column(var = "WBGeneID"), 
#           file = "../03_output/res_embryoGFPplus_vs_embryoGFPminus_apeglm.csv")
# 
# write_csv(as.data.frame(res_L1GFPplus_vs_L1GFPminus_apeglm) %>% rownames_to_column(var = "WBGeneID"), 
#           file = "../03_output/res_L1GFPplus_vs_L1GFPminus_apeglm.csv")
# 
# write_csv(as.data.frame(res_L3GFPplus_vs_L3GFPminus_apeglm) %>% rownames_to_column(var = "WBGeneID"), 
#           file = "../03_output/res_L3GFPplus_vs_L3GFPminus_apeglm.csv")
```

```{r fig.width=6, fig.height=2}
par(mfrow=c(1,3),mar=c(2,2,1,1))
ylim <- c(-10,10)
# drawLines <- function() abline(h=c(-2,2),col="dodgerblue",lwd=2)
sig = 0.01
plotMA(res_embryoGFPplus_vs_embryoGFPminus_ashr, ylim=ylim, main = "Embryo GFP+ vs GFP-", alpha = sig)
plotMA(res_L1GFPplus_vs_L1GFPminus_ashr, ylim=ylim, main = "L1 GFP+ vs GFP-", alpha = sig)
plotMA(res_L3GFPplus_vs_L3GFPminus_ashr, ylim=ylim, main = "L3 GFP+ vs GFP-", alpha = sig)
```

```{r fig.width=9, fig.height=4}
plotCounts(dds, "WBGene00001578", intgroup = "group", main = "ges-1 read counts")
```


```{r fig.width=9, fig.height=4}
plotCounts(dds, "WBGene00001578", intgroup = "group", returnData = TRUE) %>% 
  separate(group, sep = "(?<=embryo)|(?<=L1)|(?<=L3)", into = c("stage", "sample"), remove = FALSE) %>%
  ggplot(aes(x = sample, y = log10(count))) +
  geom_boxplot() +
  geom_point() +
  facet_grid(~stage) +
  ggtitle("ges-1 read counts") +
  theme_classic()
```
```{r fig.width=9, fig.height=4}
plotCounts(dds, "WBGene00001250", intgroup = "group", main = "elt-2 read counts")

plotCounts(dds, "WBGene00001250", intgroup = "group", returnData = TRUE) %>% 
  separate(group, sep = "(?<=embryo)|(?<=L1)|(?<=L3)", into = c("stage", "sample"), remove = FALSE) %>%
  ggplot(aes(x = sample, y = log10(count))) +
  geom_boxplot() +
  geom_point() +
  facet_grid(~stage) +
  ggtitle("elt-2 read counts") +
  theme_classic()
```
# Annotate and quantify tissue specific genes

```{r}
tissue_specific_genes <- read_csv(file = "../../01_tissue_specific_genes/03_output/tissue_specific_genes_220202.csv", show_col_types = FALSE)
```
```{r}
tissue_annotated_MA(res_embryoGFPplus_vs_embryoGFPminus_ashr, embryo_alt_hyp_res_df)
```
```{r}
tissue_annotated_MA(res_L1GFPplus_vs_L1GFPminus_ashr, L1_alt_hyp_res_df)
```

```{r}
tissue_annotated_MA(res_L3GFPplus_vs_L3GFPminus_ashr, L3_alt_hyp_res_df)
```



```{r fig.height=5, fig.width=6.5}

tissue_gene_quant(embryo_alt_hyp_res_df)
```
```{r fig.height=5, fig.width=6.5}
tissue_gene_quant(L1_alt_hyp_res_df)
```

```{r fig.height=5, fig.width=6.5}
tissue_gene_quant(L3_alt_hyp_res_df)
```
# Between-stage GFP+ comparisons

```{r}
res_embryoGFPplus_vs_L1GFPplus <- results(dds, contrast = c("group", "embryoGFPplus", "L1GFPplus"))
res_embryoGFPplus_vs_L3GFPplus <- results(dds, contrast = c("group", "embryoGFPplus", "L3GFPplus"))
res_L3GFPplus_vs_L1GFPplus <- results(dds, contrast = c("group", "L1GFPplus", "L3GFPplus"))

res_embryoGFPplus_vs_L1GFPplus_ashr <- lfcShrink(dds, contrast = c("group", "embryoGFPplus", "L1GFPplus"), type = "ashr")
res_embryoGFPplus_vs_L3GFPplus_ashr <- lfcShrink(dds, contrast = c("group", "embryoGFPplus", "L3GFPplus"), type = "ashr")
res_L3GFPplus_vs_L1GFPplus_ashr <- lfcShrink(dds, contrast = c("group", "L3GFPplus", "L1GFPplus"), type = "ashr")
```
```{r}
par(mfrow=c(1,3),mar=c(2,2,1,1))
ylim <- c(-15,15)
# drawLines <- function() abline(h=c(-2,2),col="dodgerblue",lwd=2)
# sig = 0.01
plotMA(res_embryoGFPplus_vs_L1GFPplus, ylim = ylim, main = "embryo_vs_L1", alpha = 0.01)
plotMA(res_embryoGFPplus_vs_L3GFPplus, ylim = ylim, main = "embryo_vs_L3", alpha = 0.01)
plotMA(res_L3GFPplus_vs_L1GFPplus, ylim = ylim, main = "L3_vs_L1", alpha = 0.01)
```

```{r}
par(mfrow=c(1,3),mar=c(2,2,1,1))
ylim <- c(-15,15)
# drawLines <- function() abline(h=c(-2,2),col="dodgerblue",lwd=2)
# sig = 0.01
plotMA(res_embryoGFPplus_vs_L1GFPplus_ashr, ylim = ylim, main = "embryo_vs_L1", alpha = 0.01)
plotMA(res_embryoGFPplus_vs_L3GFPplus_ashr, ylim = ylim, main = "embryo_vs_L3", alpha = 0.01)
plotMA(res_L3GFPplus_vs_L1GFPplus_ashr, ylim = ylim, main = "L3_vs_L1", alpha = 0.01)
```
```{r}
res_between_stage_GFP_df <- data.frame(as.data.frame(res_embryoGFPplus_vs_L1GFPplus_ashr), comparison = "embryo_vs_L1") %>%
  bind_rows(data.frame(as.data.frame(res_embryoGFPplus_vs_L3GFPplus_ashr), comparison = "embryo_vs_L3")) %>%
  bind_rows(data.frame(as.data.frame(res_L3GFPplus_vs_L1GFPplus_ashr), comparison = "L3_vs_L1")) %>% rownames_to_column(var = "WBGeneID")

res_between_stage_GFP_df %>% replace_na(list(padj = 1)) %>% filter(padj <= 0.01) %>%
  # filter(log2FoldChange >= 1 | log2FoldChange <= -1 , padj <= 0.01) %>%
  ggplot(aes(x = log10(baseMean), y = log2FoldChange)) +
  geom_point(shape = 16, alpha = 0.1, stroke = 0, size = 1, color = "red") +
  geom_point(data = res_between_stage_GFP_df %>% replace_na(list(padj = 1)) %>% filter(padj > 0.01),shape = 16, alpha = 0.1, stroke = 0, size = 1, color = "black") +
  ylim(-10,10) +
  facet_grid(~comparison) +
  theme_classic()
```
```{r}
par(mfrow=c(3,3),mar=c(2,2,1,1))
ylim <- c(-15,15)
plotMA(results(dds, contrast = c("group", "embryoGFPplus", "L1GFPplus"), lfcThreshold=thresh, altHypothesis = "greater", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "embryoGFPplus", "L1GFPplus"), lfcThreshold=thresh, altHypothesis = "less", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "embryoGFPplus", "L1GFPplus"), lfcThreshold=thresh, altHypothesis = "lessAbs", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "embryoGFPplus", "L3GFPplus"), lfcThreshold=thresh, altHypothesis = "greater", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "embryoGFPplus", "L3GFPplus"), lfcThreshold=thresh, altHypothesis = "less", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "embryoGFPplus", "L3GFPplus"), lfcThreshold=thresh, altHypothesis = "lessAbs", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "L3GFPplus", "L1GFPplus"), lfcThreshold=thresh, altHypothesis = "greater", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "L3GFPplus", "L1GFPplus"), lfcThreshold=thresh, altHypothesis = "less", alpha = sig), ylim= ylim)
plotMA(results(dds, contrast = c("group", "L3GFPplus", "L1GFPplus"), lfcThreshold=thresh, altHypothesis = "lessAbs", alpha = sig), ylim= ylim)
```
```{r}
# plotPCA(all_samples_rld, intgroup = "group")
```
Fix this to reflect new altHypothesis approach


<!-- ```{r} -->
<!-- tissue_annotated_MA(res_embryoGFPplus_vs_L1GFPplus_ashr) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- tissue_gene_quant(res_embryoGFPplus_vs_L1GFPplus_ashr) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- tissue_annotated_MA(res_L3GFPplus_vs_L1GFPplus_ashr) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- tissue_gene_quant(res_L3GFPplus_vs_L1GFPplus_ashr) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- tissue_annotated_MA(res_embryoGFPplus_vs_L3GFPplus_ashr) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- tissue_gene_quant(res_L3GFPplus_vs_L1GFPplus) -->
<!-- ``` -->


# Export rlog counts

```{r}
all_samples_rld <- rlog(dds)
# write_rds(all_samples_rld, file = "../03_output/rlog_counts/all_samples_rlog_counts.rds")
# write_tsv(as.data.frame(assay(all_samples_rld)) %>% rownames_to_column(var = "WBGeneID"), file = "../03_output/rlog_counts/all_samples_rlog_counts.tsv")
# write_tsv(as.data.frame(assay(all_samples_rld)) %>% rownames_to_column(var = "WBGeneID") %>% select(WBGeneID, contains("GFPplus")), file = "../03_output/rlog_counts/GFPplus_samples_rlog_counts.tsv")
all_samples_rld <- read_rds(file = "../03_output/all_samples_rlog_counts.rds")
all_samples_rld_df <- read_tsv(file = "../03_output/all_samples_rlog_counts.tsv")

head(all_samples_rld_df)
```

# Export intestine enrichment assignment

```{r}
embryo_intestine_gene_cats <- embryo_alt_hyp_res_df %>% 
  drop_na(padj) %>%
  filter(label == "GFPplus_vs_GFPminus", padj < sig) %>% 
  select(WBGeneID, altHyp = "type") %>% 
  mutate(intestine_expression = case_when(
    altHyp == "greater" ~ "enriched",
    altHyp == "less" ~ "depleted",
    altHyp == "lessAbs" ~ "equal")) %>% 
  mutate(intestine_expression = fct_relevel(intestine_expression, c("enriched", "equal", "depleted")))


L1_intestine_gene_cats <- L1_alt_hyp_res_df %>% 
  filter(label == "GFPplus_vs_GFPminus", padj < sig) %>% 
  select(WBGeneID, altHyp = "type") %>% 
  mutate(intestine_expression = case_when(
    altHyp == "greater" ~ "enriched",
    altHyp == "less" ~ "depleted",
    altHyp == "lessAbs" ~ "equal"))%>% 
  mutate(intestine_expression = fct_relevel(intestine_expression, c("enriched", "equal", "depleted")))

L3_intestine_gene_cats <- L3_alt_hyp_res_df %>% 
  filter(label == "GFPplus_vs_GFPminus", padj < sig) %>% 
  select(WBGeneID, altHyp = "type") %>% 
  mutate(intestine_expression = case_when(
    altHyp == "greater" ~ "enriched",
    altHyp == "less" ~ "depleted",
    altHyp == "lessAbs" ~ "equal"))%>% 
  mutate(intestine_expression = fct_relevel(intestine_expression, c("enriched", "equal", "depleted")))

# write_csv(x = embryo_intestine_gene_cats, file = "../03_output/intestine_gene_categories/embryo_intestine_gene_categories.csv")
# 
# write_csv(x = L1_intestine_gene_cats, file = "../03_output/intestine_gene_categories/L1_intestine_gene_categories.csv")
# 
# write_csv(x = L3_intestine_gene_cats, file = "../03_output/intestine_gene_categories/L3_intestine_gene_categories.csv")
```


# Average embryo GFP+ sample reads

```{r}
thresh = 1
sig = 0.01
embryo_rlog_status_df <- all_samples_rld_df %>% 
  select(WBGeneID, embryo_GFPplus_rep1, embryo_GFPplus_rep2, embryo_GFPplus_rep3) %>% 
  pivot_longer(cols = embryo_GFPplus_rep1:embryo_GFPplus_rep3, values_to = "rlog_counts") %>%
  separate(name, sep = "_", into = c("stage", "sample", "rep")) %>%
  group_by(WBGeneID) %>%
  summarise(mean.rlog.counts = mean(rlog_counts), var.rlog.counts = var(rlog_counts)) %>%
  left_join(embryo_alt_hyp_res_df %>% filter(label == "GFPplus_vs_GFPminus") %>% select(WBGeneID, type, isDE), by = "WBGeneID")

head(embryo_rlog_status_df)
# write_csv(embryo_rlog_status_df, file = "../03_output/embryo_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
# embryo_rlog_status_df <- read_csv(file = "../03_output/embryo_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
```

# Average embryo GFP+ sample reads

Make function
```{r}
thresh = 1
sig = 0.01

rlog_status <- function(stage, res, hyp_df){
all_samples_rld_df %>% 
  select(WBGeneID, contains(paste(stage,"GFPplus", sep = "_"))) %>% 
  pivot_longer(cols = contains("GFPplus"), values_to = "rlog_counts") %>%
  separate(name, sep = "_", into = c("stage", "sample", "rep")) %>%
  group_by(WBGeneID) %>%
  summarise(mean.rlog.counts = mean(rlog_counts), var.rlog.counts = var(rlog_counts)) %>%
  left_join(hyp_df %>% filter(label == "GFPplus_vs_GFPminus") %>% select(WBGeneID, type, isDE), by = "WBGeneID")
}
```


```{r}
embryo_rlog_status_df <- rlog_status(stage = "embryo", res = res_embryoGFPplus_vs_embryoGFPminus, hyp_df = embryo_alt_hyp_res_df)
# head(embryo_rlog_status_df)
# write_csv(embryo_rlog_status_df, file = "../03_output/embryo_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
# embryo_rlog_status_df <- read_csv(file = "../03_output/embryo_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
```
```{r}
L1_rlog_status_df <- rlog_status(stage = "L1", res = res_L1GFPplus_vs_L1_GFPminus, hyp_df = L1_alt_hyp_res_df)
# write_csv(L1_rlog_status_df, file = "../03_output/L1_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
# L1_rlog_status_df <- read_csv(file = "../03_output/L1_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
# head(L1_rlog_status_df)
```
```{r}
L3_rlog_status_df <- rlog_status(stage = "L3", res = res_L3GFPplus_vs_L3_GFPminus, hyp_df = L3_alt_hyp_res_df)
# write_csv(L3_rlog_status_df, file = "../03_output/L3_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
# L3_rlog_status_df <- read_csv(file = "../03_output/L3_GFPplus_rlog_counts_status_df.csv", col_names = TRUE)
# head(L3_rlog_status_df)
```
# Intesinte expression across development

# UpSet Plot
```{r}


intestine_enriched_genes <- data.frame(embryo_rlog_status_df, stage = "embryo") %>% bind_rows(data.frame(L1_rlog_status_df, stage = "L1"), data.frame(L3_rlog_status_df, stage = "L3")) %>%filter(type == "greater", isDE == TRUE) %>% select(WBGeneID, stage)

stage_list<- list(embryo = filter(intestine_enriched_genes, stage == "embryo")$WBGeneID,
     L1 = filter(intestine_enriched_genes, stage == "L1")$WBGeneID,
     L3 = filter(intestine_enriched_genes, stage == "L3")$WBGeneID)
comb_mat <-make_comb_mat(stage_list)
UpSet(comb_mat)
```
```{r}
comb_size(comb_mat)
```

# heatmap

```{r}


gfpplus_matrix <- all_samples_rld_df %>% select(WBGeneID, contains("GFPplus")) %>% filter(WBGeneID %in% intestine_enriched_genes$WBGeneID) %>% column_to_rownames(var = "WBGeneID") %>% as.matrix()

nrow(gfpplus_matrix)
```

Mean and SD thresholds
```{r}
sd_thresh <- 1
mean_thresh <- 6
```


```{r fig.width=3, fig.height=3}
hist(rowMeans(gfpplus_matrix), main = "GFPplus rlog row means")
abline(v = mean_thresh, col = "red", lty = 2)
```


```{r fig.width=3, fig.height=3}
hist(rowSds(gfpplus_matrix), main = "GFPplus rlog row SD")
abline(v = sd_thresh, col = "red", lty = 2)
```
```{r fig.width=3, fig.height=3}
plot(rowMeans(gfpplus_matrix), rowSds(gfpplus_matrix), pch = 20, cex = 0.1)
abline(h = sd_thresh, v = mean_thresh, col = "red", lty = 2)
# abline(, col = "red", lty = 2)
```


```{r}
gfpplus_matrix_row_zscore <- (gfpplus_matrix - rowMeans(gfpplus_matrix))/rowSds(gfpplus_matrix)
```

```{r fig.width=5, fig.height=5}
pheatmap(gfpplus_matrix_row_zscore, 
         cluster_cols = FALSE, 
         show_rownames = FALSE, 
         # cutree_rows = 5,
         main = paste0("Dynamics of intestine enriched genes\n unthresholded\n # genes = ", nrow(gfpplus_matrix_row_zscore), sep = ""))
```


```{r fig.width=5, fig.height=5}
pheatmap(gfpplus_matrix_row_zscore[rowSds(gfpplus_matrix) >= sd_thresh & rowMeans(gfpplus_matrix) >= mean_thresh,], 
         cluster_cols = TRUE, 
         show_rownames = FALSE, 
         # cutree_rows = 5,
         main = paste0("Dynamics of intestine enriched genes\n thresholds: row mean >= ", mean_thresh, ", row SD >= ", sd_thresh,
                       "\n # genes = ", nrow(gfpplus_matrix_row_zscore[rowSds(gfpplus_matrix) >= sd_thresh & rowMeans(gfpplus_matrix) >= mean_thresh,])))
```
# Heatmap of transcription factors

```{r fig.width=5, fig.height=4}
wtf3 <- read_table(file = "../../../David/01_promoters/01_input/wtf3.wbid", col_names = FALSE)$X1

gfpplus_matrix_row_zscore_TFs <- gfpplus_matrix_row_zscore[rownames(gfpplus_matrix_row_zscore) %in% wtf3,]
pheatmap(gfpplus_matrix_row_zscore_TFs ,
         cluster_cols = FALSE, 
         show_rownames = FALSE,
         main = paste0("Dynamics of intestine enrcihed transcription factors\n unthresholded\n # genes =", 
                       nrow(gfpplus_matrix_row_zscore_TFs ), sep = "")
)
```
```{r fig.width=3, fig.height=3}
gfpplus_matrix_TFs <- gfpplus_matrix[rownames(gfpplus_matrix) %in% wtf3,]
plot(rowMeans(gfpplus_matrix_TFs), rowSds(gfpplus_matrix_TFs))
abline(h = 1, col = "red", lty = 2)
```
```{r fig.width=5, fig.height=4}
pheatmap(gfpplus_matrix_row_zscore_TFs[rowSds(gfpplus_matrix_TFs) >= 1,] ,
         cluster_cols = FALSE, 
         show_rownames = FALSE,
         main = paste0("Dynamics of intestine enrcihed transcription factors\n theshold: row SD >= 1\n # genes =", 
                       nrow(gfpplus_matrix_row_zscore_TFs[rowSds(gfpplus_matrix_TFs) >= 1,]), sep = "")
)
```

# Data output

```{r}
# # per stage GFP+ vs GFP- differential expression
# write_csv(res_to_df(res_embryoGFPplus_vs_embryoGFPminus), file = "../03_output/pairwise_DE_results/res_embryoGFPplus_vs_embryoGFPminus.csv")
# write_csv(res_to_df(res_L1GFPplus_vs_L1_GFPminus), file = "../03_output/pairwise_DE_results/res_L1GFPplus_vs_L1GFPminus.csv")
# write_csv(res_to_df(res_L3GFPplus_vs_L3_GFPminus), file = "../03_output/pairwise_DE_results/res_L3GFPplus_vs_L3GFPminus.csv")
# 
# # per stage GFP+ vs GFP- differential expression with log2FC shrink (visualization, ranking)
# write_csv(res_to_df(res_embryoGFPplus_vs_embryoGFPminus_ashr), file = "../03_output/pairwise_shrunk_DE_results/res_embryoGFPplus_vs_embryoGFPminus_ashr_shrunk.csv")
# write_csv(res_to_df(res_L1GFPplus_vs_L1GFPminus_ashr), file = "../03_output/pairwise_shrunk_DE_results/res_L1GFPplus_vs_L1GFPminus_ashr_shrunk.csv")
# write_csv(res_to_df(res_L3GFPplus_vs_L3GFPminus_ashr), file = "../03_output/pairwise_shrunk_DE_results/res_L3GFPplus_vs_L3GFPminus_ashr_shrunk.csv")
# 
# # per stage enrichment annotation
# write_csv(embryo_alt_hyp_res_df %>% filter(label == "GFPplus_vs_GFPminus") %>% select(WBGeneID:type, isDE), file = "../03_output/embryo_GFPplus_vs_GFPminus_alt_hyp_res.csv")
# write_csv(L1_alt_hyp_res_df %>% filter(label == "GFPplus_vs_GFPminus") %>% select(WBGeneID:type, isDE), file = "../03_output/L1_GFPplus_vs_GFPminus_alt_hyp_res.csv")
# write_csv(L3_alt_hyp_res_df %>% filter(label == "GFPplus_vs_GFPminus") %>% select(WBGeneID:type, isDE), file = "../03_output/L3_GFPplus_vs_GFPminus_alt_hyp_res.csv")
```

# Plot output

```{r}
res_embryoGFPplus_vs_embryoGFPminus_ashr <- read_csv(file = "../03_output/pairwise_shrunk_DE_results/res_L3GFPplus_vs_L3GFPminus_ashr_shrunk.csv")
embryo_intestine_gene_cats <- read_csv(file = "../03_output/intestine_gene_categories/L3_intestine_gene_categories.csv")
res_embryoGFPplus_vs_embryoGFPminus_ashr %>% left_join(embryo_intestine_gene_cats, by = "WBGeneID") %>% drop_na(intestine_expression) %>% filter(intestine_expression == "enriched") %>%
  ggplot(aes(x =  log10(baseMean), y =log2FoldChange)) +
  geom_point(data = res_embryoGFPplus_vs_embryoGFPminus_ashr %>% filter(!(WBGeneID %in%embryo_intestine_gene_cats$WBGeneID)),
             shape = 16, alpha = 0.5, stroke = 0, size = 1, color = "grey") +
  geom_point(shape = 16, alpha = 0.5, stroke = 0, size = 1, aes(color = intestine_expression)) +
  theme_bw() +
  labs(x = "log10(mean RNA abundance)",
       y = "log2(embryo GFP+ reads/embryo GFP- reads)")
ggsave(filename = "../03_output/plots/Intestine_Expression_Category/L3_GFPplus_vs_GFPminus_shrunk_MAplot_enriched_only.jpg", width = 5, height = 3)
  
res_to_df(res_embryoGFPplus_vs_embryoGFPminus_ashr) %>% left_join(embryo_intestine_gene_cats, by = "WBGeneID") %>% drop_na(intestine_expression) %>% group_by(intestine_expression) %>% summarise(genes = n()) %>% ggplot(aes(x = intestine_expression, y = genes, fill = intestine_expression, label = genes)) +
  geom_bar(stat = "identity") +
  geom_text(vjust = -0.25) +
  theme_bw()

```
```{r}
single_MA_plot <- function(in_res, in_cats){
  res_to_df(in_res) %>% left_join(in_cats, by = "WBGeneID") %>% drop_na(intestine_expression) %>%
    filter(log2FoldChange > -20) %>%
  ggplot(aes(x =  log10(baseMean), y =log2FoldChange)) +
  geom_point(data = res_to_df(in_res) %>% filter(!(WBGeneID %in% in_cats$WBGeneID)),
             shape = 16, alpha = 0.5, stroke = 0, size = 1, color = "grey") +
  geom_point(shape = 16, alpha = 0.5, stroke = 0, size = 1, aes(color = intestine_expression)) +
  theme_bw() +
  labs(x = "log10(mean RNA abundance)",
       y = "log2(GFP+ reads/GFP- reads)",
       title = paste("data: ", deparse(substitute(in_res)), sep = ""))
}

single_cat_bar <- function(in_res, in_cats){
  res_to_df(in_res) %>% left_join(in_cats, by = "WBGeneID") %>% drop_na(intestine_expression) %>% group_by(intestine_expression) %>% summarise(genes = n()) %>% ggplot(aes(x = intestine_expression, y = genes, fill = intestine_expression, label = genes)) +
  geom_bar(stat = "identity") +
  geom_text(vjust = -0.25) +
  ggtitle(paste("data: ", deparse(substitute(in_cats)), sep = "")) +
  theme_bw()
}
```

```{r}
embryo_shrunk_MA <- single_MA_plot(res_embryoGFPplus_vs_embryoGFPminus_ashr, embryo_intestine_gene_cats)
embryo_shrunk_MA
# ggsave(plot = embryo_shrunk_MA, file = "../03_output/plots/Intestine_Expression_Category/embryo_GFPplus_vs_GFPminus_shrunk_MAplot.pdf", width = 5, height = 3)

embryo_category_bar <- single_cat_bar(res_embryoGFPplus_vs_embryoGFPminus_ashr, embryo_intestine_gene_cats)
embryo_category_bar
# ggsave(plot = embryo_category_bar, file = "../03_output/plots/Intestine_Expression_Category/embryo_GFPplus_vs_GFPminus_Expression_Category_barplot.pdf", width = 5, height = 3)
```

```{r}
L1_shrunk_MA <- single_MA_plot(res_L1GFPplus_vs_L1GFPminus_ashr, L1_intestine_gene_cats)
L1_shrunk_MA
# ggsave(plot = L1_shrunk_MA, file = "../03_output/plots/Intestine_Expression_Category/L1_GFPplus_vs_GFPminus_shrunk_MAplot.pdf", width = 5, height = 3)

L1_category_bar <- single_cat_bar(res_L1GFPplus_vs_L1GFPminus_ashr, L1_intestine_gene_cats)
L1_category_bar
# ggsave(plot = L1_category_bar, file = "../03_output/plots/Intestine_Expression_Category/L1_GFPplus_vs_GFPminus_Expression_Category_barplot.pdf", width = 5, height = 3)
```

```{r}
L3_shrunk_MA <- single_MA_plot(res_L3GFPplus_vs_L3GFPminus_ashr, L3_intestine_gene_cats)
L3_shrunk_MA
# ggsave(plot = L3_shrunk_MA, file = "../03_output/plots/Intestine_Expression_Category/L3_GFPplus_vs_GFPminus_shrunk_MAplot.pdf", width = 5, height = 3)

L3_category_bar <- single_cat_bar(res_L3GFPplus_vs_L3GFPminus_ashr, L3_intestine_gene_cats)
L3_category_bar
# ggsave(plot = L3_category_bar, file = "../03_output/plots/Intestine_Expression_Category/L3_GFPplus_vs_GFPminus_Expression_Category_barplot.pdf", width = 5, height = 3)
```

# Alternative Hypothesis output

```{r}
thresh <- 1
sig <- 0.01

res_embryoGFP_alHyp_greater <- results(dds, contrast = c("group", "embryoGFPplus", "embryoGFPminus"), lfcThreshold=thresh, altHypothesis = "greater", alpha = sig)
write_csv(x = res_to_df(res_embryoGFP_alHyp_greater), file = "../03_output/res_embryoGFP_alHyp_greater.csv")

res_L1GFP_alHyp_greater <- results(dds, contrast = c("group", "L1GFPplus", "L1GFPminus"), lfcThreshold=thresh, altHypothesis = "greater", alpha = sig)
write_csv(x = res_to_df(res_L1GFP_alHyp_greater), file = "../03_output/res_L1GFP_alHyp_greater.csv")

res_L3GFP_alHyp_greater <- results(dds, contrast = c("group", "L3GFPplus", "L3GFPminus"), lfcThreshold=thresh, altHypothesis = "greater", alpha = sig)
write_csv(x = res_to_df(res_L3GFP_alHyp_greater), file = "../03_output/res_L3GFP_alHyp_greater.csv")


```


